\abstract{

Network function (NF) migration alongside (and possibly because of)
routing policy updates is a delicate task, making it difficult to
ensure that all traffic is processed by its required network
functions, in order. Achieving traffic redistribution while
ensuring correct processing of all packets requires an efficient network 
forwarding-state update and careful coordination between routing-policy 
change and NF migration. To achieve consistent network updates, in this dissertation,
we propose a new method that is inspired by \textit{causal consistency}, a consistency model for
shared-memory systems.  We propose and analyze a property called \textit{suffix causal consistency} (SCC) as an interpretation of causal consistency for rule updates in an SDN network. We design an algorithm implementing this property and formally verify the correctness of this algorithm using model checking. Our evaluation results show that SCC
provides greater efficiency than competing consistent-update
alternatives while offering consistency that is strong enough to
ensure high-level routing properties (e.g., black-hole freedom).

To coordinate routing-policy update with NF migration, we propose a design called \sysname for
interleaving these tasks to achieve more efficient
completion of both while ensuring complete processing of traffic by
the required sequences of NFs. Our technique works with any route-update protocol that implements a property we call \textit{relaxed waypoint
correctness}, which includes our SCC algorithm and many consistent-update protocols.
We also provide a route-update protocol that is customized to
achieve relaxed waypoint correctness without conforming to
conventional ``consistent update'' semantics, as typically defined for
such protocols.  We confirm the sufficiency of relaxed waypoint
correctness using model checking, and the implementation demonstrates the efficiency and efficacy of \sysname.

}
